### 2022-12-20 source/abstract.tex
polish

% \cleardoublepage
% \pdfbookmark{Zusammenfassung}{abstract_DE}

% { % need this extra grouping to avoid carry-over effects
% \begin{otherlanguage}{ngerman}
% \enlargethispage{4\baselineskip}
% \begin{center}\abstractnamefont
%   Lorem ipsum dolor sit amet, consectetur adipiscing elit, sed do eiusmod tempor
%   incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis
%   nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat.
%   Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore
%   eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt
%   in culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum dolor
%   sit amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut
%   labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud
%   exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis
%   aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu
%   fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in
%   culpa qui officia deserunt mollit anim id est laborum. Lorem ipsum dolor sit
%   amet, consectetur adipiscing elit, sed do eiusmod tempor incididunt ut labore
%   et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation
%   ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor
%   in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla
%   pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui
%   officia deserunt mollit anim id est laborum.

% \end{center}
% \end{otherlanguage}
% }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

% The credibility of a scientific finding is based on the fact that it can be
% replicated when the same study is repeated with new subjects.
% An important aspect of the credibility of a scientific finding is that the
% finding can be replicated when the same study is repeated with new subjects.
% An important aspect of the credibility of a scientific finding is its
% replicability -- the ability to produce similar results when the same study is
% repeated with new subjects.
An important aspect of the credibility of scientific findings is their
replicability -- the ability that a similar finding can be obtained when the
same study is repeated with new subjects. However, various failures to replicate
major scientific findings in the social and life sciences indicate that
replicability is often lower than expected. This ``replication crisis'' has led
to several methodological reforms in the past decade, an increased conduct of
replication studies being one of them. Despite this increase in replication
studies, there is no consensus on a fundamental question: How should replication
success be defined statistically?
% Various methods already exist and various new ones have been proposed in
% response to the crisis. This thesis centres around a proposal based on the
% reverse-Bayes approach. The key idea is to reverse the traditional ``forward''
% use of Bayes' theorem (``prior + likelihood $\rightarrow$ posterior''),
% starting instead with a pre-specified posterior and deducing the corresponding
% prior (``posterior + likelihood $\rightarrow$
% prior''). % As such, the reverse-Bayes approach naturally fits to the replication
% % setting as it enables the mathematical formalization of scepticism regarding the
% % finding from the original study.
% This allows us to challenge the original finding by determining a sceptical
% prior for the underlying effect size such that the resulting posterior is no
% longer convincing. This prior then represents the position of a sceptic who
% remains unconvinced by the original study. Whether or not the sceptics'
% position is justified can then be assessed in light of the new data from the
% replication study. The larger the conflict between the replication data and
% the sceptic, the larger the degree of replication success. This procedure can
% be summarized in a single quantitative measure of replication success, termed
% the sceptical $p$-value.

Possible answers to this question are given in the first and largest part of
this thesis. It consists of three extensions to a recently proposed
reverse-Bayes method for quantifying replication success. The key idea of the
method is to challenge the data from the original study with a sceptical prior
distribution so that the resulting posterior distribution no longer indicates
evidence for an effect. The goal of the replication study is then to show that
the sceptical prior is unrealistic, and replication success is achieved if there
is sufficient conflict between the prior and the replication data. The procedure
can be summarized in a single quantitative measure, termed the sceptical
$p$-value.
% extension 1
The first extension recalibrates the sceptical $p$-value so that replication
success takes effect size more appropriately into account. Specifically, the
recalibration is chosen such that for original studies which were borderline
significant, replication success can only be achieved if the effect estimate
from the replication study is larger than the effect estimate from the original
study. We find that the recalibrated sceptical $p$-value also has good
frequentist properties comparable to the standard method used in practice.
% extension 2
The second extension replaces tail probabilities with Bayes factors as measures
of evidence. The procedure can again be summarized in a single measure called
the sceptical Bayes factor.
% The sceptical prior is now determined such that the original finding is no
% longer convincing in terms of a Bayes factor. In contrast to tail
% probabilities,
% We find that the extended method combines two notions of replicability: It
% requires that original and replication study independently provide evidence for
% an effect and ensures that their effect estimates are compatible.
The sceptical Bayes factor has similar properties as the sceptical $p$-value but
it avoids a statistical paradox which the sceptical $p$-value suffers from -- in
some situations the sceptical $p$-value may indicate replication success even
though the effect estimate from the replication is substantially smaller than
the effect estimate from the original study. % , which cannot happen with the sceptical
% Bayes factor.
% Bayes factors have a more
% natural interpretation and allow for direct quantification of evidence for one
% hypothesis versus another. As with the sceptical $p$-value, the procedure leads
% to a single measure quantifying the degree of replication success called the
% sceptical Bayes factor.
The third extension provides a framework for Bayesian design of replication
studies which allows for combining data from the original study with external
knowledge. This approach can lead to potentially more efficient designs compared
to classical approaches, and it can be used for replication design based on both
the sceptical $p$-value and the sceptical Bayes factor.
% External knowledge, such as data from other studies, is ubiquitous in the
% replication setting and taking it into account can lead to potentially more
% efficient designs compared to classical approaches.

>>>>

@@

An important aspect of the credibility of scientific findings is their
replicability -- the ability that a similar finding can be obtained when a study
is repeated with new subjects. However, various failures to replicate major
scientific findings in the social and life sciences indicate that replicability
is often lower than expected. This ``replication crisis'' has led to several
methodological reforms in the past decade, an increased conduct of replication
studies being one of them. Despite this increase in replication studies, there
is no consensus on a fundamental question: How should replication success be
defined statistically?

Possible answers to this question are given in the first and largest part of
this thesis. It consists of three extensions to a recently proposed
reverse-Bayes method for quantifying replication success. The methods' key idea
is to challenge the data from the original study with a sceptical prior
distribution for the effect size so that the resulting posterior distribution no
longer indicates evidence for an effect. The goal of the replication study is
then to show that the sceptical prior is unrealistic, and replication success is
achieved if there is sufficient conflict between the prior and the replication
data. The procedure can be summarized in a single quantitative measure, termed
the sceptical $p$-value.
% extension 1
The first extension recalibrates the sceptical $p$-value so that replication
success takes effect size more appropriately into account. Specifically, the
recalibration is chosen such that for original studies which were borderline
significant, replication success can only be achieved if the effect estimate
from the replication study is larger than the effect estimate from the original
study. We find that the recalibrated sceptical $p$-value also has good
frequentist properties comparable to the standard method used in practice.
% extension 2
The second extension replaces tail probabilities with Bayes factors as measures
of evidence. The procedure can again be summarized in a single measure called
the sceptical Bayes factor. The sceptical Bayes factor has similar properties as
the sceptical $p$-value but it avoids a statistical paradox which the sceptical
$p$-value suffers from -- in some situations the sceptical $p$-value may
indicate replication success even though the effect estimate from the
replication is substantially smaller than the effect estimate from the original
study.
% extension 3
The third extension provides a framework for Bayesian design of replication
studies which allows for combining data from the original study with external
knowledge. This approach can lead to potentially more efficient designs compared
to classical approaches, and it can be used for replication design based on both
the sceptical $p$-value and the sceptical Bayes factor.

### 2022-12-20 source/intro.Rnw
polish

Despite their importance, the traditional academic system has made it
unattractive for researchers to conduct replication studies as publications,
citations, and grant money are typically easier to acquire by conducting novel
research. This is because until recently, most scientific journals hardly
published any replication studies since the emphasis was on ``innovation''
rather than on ``repetition'' \citep{Makel2012, Martin2017, Coiera2021}. As a
result, researchers often had no other choice than to focus their efforts on
producing novel and eye-catching research results in order to build successful
careers \citep{Binswanger2014, Moher2018}.

@@

Most methods for analyzing replication studies can be formulated in terms of
study-level summary statistics as shown in Table~\ref{tab:meta}. All of these
are routinely reported in research articles, and if one of them is missing they
can typically be calculated from the others. Using summary statistics is also
often the only possible way for conducting the replication success analysis as
the raw data from the original study may not be available to the researchers
conducting the replication study. The most important statistics are the effect
estimates $\that_{o}$ and $\that_{r}$. They provide an estimate of the
underlying effect size $\theta$ which quantifies the true effect or association
of an intervention/exposure with an outcome variable. Typical effect sizes are
mean differences and correlations (for continuous outcomes), odds ratios (for
binary outcomes), or hazard ratios (for time-to-event outcomes). Depending on
the type of effect size, a transformation might be required so that the
assumption of approximately normally distributed effect estimates around the
unknown effect size $\theta$ (for large enough sample sizes) is justified. This
could be, for instance, the Fisher $z$-transformation for correlations or the
log-transformation for odds/hazard ratios \citep[chapter 11]{Cooper2019}. The
associated standard errors $\sigma_{o}$ and $\sigma_{r}$ represent the
statistical uncertainty of the estimates. Typically, the available standard
errors are only estimates of the true standard errors. However, in most parts of
this thesis it will be assumed that $\sigma_{o}$ and $\sigma_{r}$ correspond to
the true standard errors. This is the same assumption as in ordinary
fixed-effect/random-effects meta-analysis, and typically reasonable if the
sample size of the studies is not too small. Under the assumption of
(asymptotic) normality, confidence intervals for $\theta$ and $p$-values for
testing the null hypothesis $H_{0} \colon \theta = 0$ can be computed as shown
in Table~\ref{tab:meta}.

@@

Table~\ref{tab:rsdef} lists commonly used criteria for replication success in
terms of the summary statistics from Table~\ref{tab:meta}. The most popular
criterion is listed first and defines replication success by simultaneous
statistical significance of original and replication study along with their
effect estimates showing the same direction. This approach is also called the
``two-trials rule'' \citep{Senn2008} in drug regulation or ``vote-counting'' in
meta-analysis \citep{Cooper2019}. The criterion can similarly be defined through
one-sided $p$-values, so that the original and replication effect estimate are
not explicitly required to show the same direction as this is already taken into
account by the one-sided $p$-values. Some replication projects
\citep{Opensc2015, Errington2021} also defined replication success via
simultaneous non-significance of original and replication study
($p_{o} > \alpha$ and $p_{r} > \alpha$). However, with this definition
``replication success'' can almost always be achieved by conducting original and
replication study with very few samples so that the $p$-values are large. The
approach is also logically questionable as it could be seen as an instance of
the ``absence of evidence fallacy'' \citep{Altman1995} meaning the failure to
find evidence against the null hypothesis is erroneously interpreted as evidence
for the null hypothesis. Meta-analytic extensions of the significance approach
define replication success by significance of a combined effect
estimate.
% \footnote{The meta-analytic approach can also be given a Bayesian
% interpretation: A flat prior distribution for the underlying effect size
% $\theta$ is updated by the data from original and replication study.
% Replication success defined by a meta-analytic $p$-value being smaller than
% $\alpha$ is then equivalent to replication success via a Bayesian posterior
% probability $\Pr(\theta > 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$,
% respectively, $\Pr(\theta < 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$,
% depending on the direction of the combined estimate.}
Typically, the assumption of a common underlying effect size is seen as
reasonable so that fixed-effect meta-analysis is used for pooling.
Random-effects meta-analysis has primarily been used if more than one
replication study of the same original study are conducted since replicators are
often interested in understanding between-replication heterogeneity \citep[\eg{}
in][]{Klein2018}. In contrast to the ordinary significance criterion, it is less
established which level $\alpha$ should be used for thresholding the
meta-analytic $p$-value. Some replication projects have used the same level as
for the significance approach \citep{Opensc2015, Camerer2016, Errington2021},
whereas others have used a smaller level as the pooled estimate is based on both
data sets \citep[\eg{} $\alpha = 0.5\%$ in][]{Camerer2018}.

>>>>

Despite their importance, the traditional academic system has made it
unattractive for researchers to conduct replication studies. This is because
until recently, most scientific journals hardly published any replication
studies as the emphasis was on ``innovation'' rather than on ``repetition''
\citep{Makel2012, Martin2017, Coiera2021}. As a result, researchers often had no
other choice than to focus their efforts on producing novel and eye-catching
research results in order to build successful careers \citep{Binswanger2014,
  Moher2018}.

@@

Most methods for analyzing replication studies can be formulated in terms of
study-level summary statistics as shown in Table~\ref{tab:meta}. All of these
are routinely reported in research articles, and if one of them is missing they
can typically be calculated from the others. Using summary statistics is also
often the only possible way for conducting the replication success analysis as
the raw data from the original study may not be available to the researchers
conducting the replication study. The effect estimates $\that_{o}$ and
$\that_{r}$ provide an estimate of the underlying effect size $\theta$ which
quantifies the true effect or association of an intervention/exposure with an
outcome variable. Typical effect sizes are mean differences and correlations
(for continuous outcomes), odds ratios (for binary outcomes), or hazard ratios
(for time-to-event outcomes). Depending on the type of effect size, a
transformation might be required so that the assumption of approximately
normally distributed effect estimates around the unknown effect size $\theta$
(for large enough sample sizes) is justified. This could be, for instance, the
Fisher $z$-transformation for correlations or the log-transformation for
odds/hazard ratios \citep[chapter 11]{Cooper2019}. The associated standard
errors $\sigma_{o}$ and $\sigma_{r}$ represent the statistical uncertainty of
the estimates. Typically, the available standard errors are only estimates of
the true standard errors. However, in most parts of this thesis it will be
assumed that $\sigma_{o}$ and $\sigma_{r}$ correspond to the true standard
errors. This is the same assumption as in ordinary fixed-effect/random-effects
meta-analysis, and typically reasonable if the sample size of the studies is not
too small. Under the assumption of (asymptotic) normality, confidence intervals
for $\theta$ and $p$-values for testing the null hypothesis
$H_{0} \colon \theta = 0$ can be computed as shown in Table~\ref{tab:meta}.

@@

Table~\ref{tab:rsdef} lists commonly used criteria for replication success in
terms of the summary statistics from Table~\ref{tab:meta}. The most popular
criterion is listed first and defines replication success by simultaneous
statistical significance of original and replication study along with their
effect estimates showing the same direction. This approach is also called the
``two-trials rule'' \citep{Senn2008} in drug regulation or ``vote-counting'' in
meta-analysis \citep{Cooper2019}. The criterion can similarly be defined through
one-sided $p$-values, so that the original and replication effect estimate are
not explicitly required to show the same direction as this is already taken into
account by the one-sided $p$-values. Some replication projects
\citep{Opensc2015, Errington2021} also defined replication success via
simultaneous non-significance of original and replication study
($p_{o} > \alpha$ and $p_{r} > \alpha$). However, with this definition
``replication success'' can almost always be achieved by conducting original and
replication study with very few samples so that the $p$-values are large. The
approach is also logically questionable as it could be seen as an instance of
the ``absence of evidence fallacy'' \citep{Altman1995} meaning the failure to
find evidence against the null hypothesis is erroneously interpreted as evidence
for the null hypothesis.

Meta-analytic extensions of the significance approach define replication success
by significance of a combined effect estimate.
% \footnote{The meta-analytic approach can also be given a Bayesian
% interpretation: A flat prior distribution for the underlying effect size
% $\theta$ is updated by the data from original and replication study.
% Replication success defined by a meta-analytic $p$-value being smaller than
% $\alpha$ is then equivalent to replication success via a Bayesian posterior
% probability $\Pr(\theta > 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$,
% respectively, $\Pr(\theta < 0 \given \that_{o}, \that_{r}) > 1 - \alpha/2$,
% depending on the direction of the combined estimate.}
Typically, the assumption of a common underlying effect size is seen as
reasonable so that fixed-effect meta-analysis is used for pooling.
Random-effects meta-analysis has primarily been used if more than one
replication study of the same original study are conducted since replicators are
often interested in understanding between-replication heterogeneity \citep[\eg{}
in][]{Klein2018}. In contrast to the ordinary significance criterion, it is less
established which level $\alpha$ should be used for thresholding the
meta-analytic $p$-value. Some replication projects have used the same level as
for the significance approach \citep{Opensc2015, Camerer2016, Errington2021},
whereas others have used a smaller level as the pooled estimate is based on both
data sets \citep[\eg{} $\alpha = 0.5\%$ in][]{Camerer2018}.
